\huge\bf{Querypoint} : Moving Backwards on Wrong Values in the Buggy Execution

### 2011-06-09 trunk/papers/fse-tool2011/sig-alternate.tex

\textit{Querypoint}, is a Firefox plugin which enhances the popular Firebug JavaScript debugger with
 a new, practical feature called \textit{lastChange}, which automatically locates the last point that a variable 
or an object property has been changed. Starting from a program halted on a breakpoint, the \textit{lastChange} solution
applies queries to the live program during re-execution, recording the call stack and limited program state each time 
the property value changes. When the program halts again on the breakpoint, the recorded information can be shown to the developer.

@@

In breakpoint debugging, developers select data to be collected by
searching through source files and setting breakpoints. To determine
where a value was set incorrectly, a developer must set
breakpoints at all possible points where the value changes. At
every breakpoint, the developer must determine if the location is in
fact related to the questionable value change then study the complex
debugger user interface and memorize values or manually collect
data. As the number of breakpoint hits increases, the process of
checking the program state, collecting data and resuming the execution
becomes cumbersome.

In log-based debugging, developers select data to be collected by
inserting statements for all points of possible change.  While in
breakpoint-based debugging, the whole program state is available to
developer, in log-based debugging, developer has to decide what data
should be collected when inserts the log statement. It is very common
that the developer has to repeat this step several times due to
insufficient collected data, or to wait a long time because too much
data is recorded. Once adequate data is collected, it still
requires analyzing and understanding. Developers usually end up in
dealing with long log files and analyzing huge amounts of collected data. 
Neither approach effectively assists the developer in finding origins to a wrong
value.

Our new functionality in
debuggers, \textit{lastChange}, locates the origin of a wrong
value by queries on the running program.
Imagine that a program execution is
paused on a breakpoint and the developer is
suspicious about the value of a variable or an object property. The
developer selects \textit{lastChange} on the value. The debugger
replays the buggy execution and collects data
when the data field changes. Once the execution reaches the same place 
(i.e., the same
breakpoint hit), it pauses the execution, analyzes the collected data
and shows the location of the last change to the developer. The
developer can also examine the program state at the located point of
execution, and continue debugging by more \textit{lastChange} queries
from that point.

Our contribution in this paper is the technique \textit{lastChange},
which locates the last place a value has changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The technique builds on existing breakpoint debugger
technology  and it does not require a special environment to
create identical, instruction by instruction, re-executions. We demonstrate the feasibility of the approach with 
\textit{Querypoint}, an implementation extending Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.
The \textit{lastChange} algorithm provides information on important program
values during the program execution without voluminous logs and without tedious
 insertion and removal of breakpoints. We believe other queries over the
 running program can be formulated to generalize this technique.

\section{Related Work}
\label{sec:relatedWork}
%whyline, firecrystal

Tools developed to enhance developers' movement on the buggy execution can be classifed in two groups: replay-based and logging-based.

%The algorithm we presented in this paper obtain information about
%the execution state logically earlier in the control flow by querying 
%the executioin state during replay using the technology of 
%conventional breakpoint debuggers.  
The \textit{lastChange} approach resembles the operational model of replay-based debugging 
and the query approach of logging-based
debugging.  Replay-based approaches capture limited data during
execution and replay the buggy execution to reach past points. In
contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution, then query the data to 
inform the developer. Replay-based
approaches impose much less runtime overhead (about two orders of
magnitudes) comparing to logging-based approaches. However, developer
has to re-execute the buggy execution several
times. \textit{lastChange} collects data on re-execution by queries
selected by developer interaction with the debugger. Therefore it has 
the selectivity of the replay-based approaches that improves performance, 
% but we have 
and the flexibilty of the queries so it does not require deterministic replay.

Among replay-based debuggers we compare to bdb \cite{Boothe} and
reverse watchpoint \cite{Maruyama}.  A bidirectional C debugger, bdb
employs a step counter to locate the requested point from the
beginning of execution. It relies on deterministic execution replay
%(i.e., the same sequence of instructions in re-execution) 
and records the results of non-deterministic system calls and re-injects them into
the program when it is replayed. It makes use of checkpoints to reduce
the time needed for re-execution. Reverse watchpoint, similar to bdb, uses
a counter to correctly locate the last write access of a selected variable in the
next execution \cite{Maruyama}.
% Reverse watchpoint is proposed by Maruyama et al., analyses the execution and moves the debugger to the
% last write access of a selected variable by re-executing the program
%from the beginning\cite{Maruyama}.  Similar to bdb it relies on deterministic
%replay and uses a counter to correctly locate a point in the next
%execution. 
The main disadvantage of these approaches is that they require 
identical, instruction by instruction, re-executions. % exactly the same executions. 
Even one instruction difference between
two executions leads to wrong results. On the other hand,
\textit{lastChange} doesn't require any special feature in the
re-execution and fits into existing debugging practice

Among logging-based approaches are \textit{omniscient} debuggers
ODB\cite{Lewis} and Unstuck\cite{Hofer}. Both
approaches keep the log history in memory and hence can only record
and store the complete history for a short period of time. These
debuggers record all the events that occur during the buggy execution
and later let the developer to navigate through the obtained execution
log. In these approaches there is no execution to resume: moving
backwards in the log can be similar to moving forwards. A more scalable 
approach to omniscient debugging has been proposed by Pothier et al. \cite{Pothier}. 
Their back-in-time debugger, TOD, addresses the space problem by storing 
execution events in a distributed database. Bhansali et al. \cite{Bhansali} attempted to address the performance overhead 
and large trace size of logging-based debugging in their time-travelling debugger, Nirvana. 
Nirvana first collects a full compressed trace of program execution and then 
simulates re-execution. Their results are quite similar to omniscient 
debuggers. Nirvana incurs about 15 to 68 times runtime overhead in 
re-simulation.

Logging-based debuggers suffer from different issues. First, the recording
step is time expensive and it should be repeated in case of changes in
program. Second, the execution log cannot fully replace the live
execution. There are other aspects of execution (e.g., program user
interface, file system, database tables, etc.) which are also
important in debugging and are not available to the developer in 
logging-based debuggers. Third, querying collected data (e.g., to restore
the program state at a certain point) may not be efficient enough for
debugging realistic programs. Comparing to logging-based debuggers our approach 
has little upfront cost and more flexibility. The developer can start debugging just 
after reproducing the bug without a capturing step.  Changing inputs or
environment settings and re-executing to investigate the bug works as
in conventional breakpoint debuggers.

A recent work by Lienhard et al.\cite{Lienhard} suggests virtual
machine level support for keeping the history of events. It replaces every
object reference with an alias object which keeps the history of
changes to the object reference. 
%In this way, when an object is
%collected by garbage collector, its track of changes (if it is not
%referenced by other aliases) will be also collected. 
Although this approach incurs less runtime overhead (7 times) in
comparison to omniscient debuggers, it adds memory
overhead.

Origin tracking of \texttt{undefined} and \texttt{null} values employing \textit{value piggybacking} technique proposed by
Bond et al. \cite{Bond}. This approach has two main limitations comparing to \textit{lastChange}.
First, it is limited to \texttt{undefined} and \texttt{null} values. Second, it does not return the last change
of a \texttt{null} variable but the first place that the \texttt{null} value is originated.

%Two new directions in logging debuggers explore more detailed use of
%the log and more effective user interface to the collected data. %logging approaches. 
%WhyLine\cite{Ko} provides visual interface to collected runtime information and lets
%developer moves on execution log using queries expressed in terms of
%the programming objects.

WhyLine\cite{Ko} stores the program user interface in
addition to the program trace and provides answers to why and why not
questions to the user. 
%Jive\cite{Czyz} depicts the history of
%execution by a sequence diagram and lets the user to query on the events
%database. 
WhyLine suffers from the requirement of gathering tracing information before its unique capabilities can be used.
We imagine that the runtime queries we use in \textit{lastChange} may be used to gather data incrementally for this kind of debugging approach.

%Querypoint debugging uses re-executions to gather infor-mation requested by the developer: the memory overhead depends on the query not the entire program. Moreover, the Lienhard et al. debugger significantly changes the virtual machine, while our approach is a generalization to conditional breakpoints and available debugger infrastructure can be adapted to support it.

%\textit{lastChange} functionality does rely on a conventional breakpoint to begin queries, a requirement not shared by full logging solutions.  Here we leverage past experience of developers, but there are also new tools [] to help with this problem in the case of graphical and event based systems.

\section{Introductory example}
\label{sec:introExample}

We illustrate the \textit{lastChange} functionality by a simple
example. The example demonstrates a buggy JavaScript code in a HTML
page (Figure~\ref{fig:js-code}). The page contains a button (line 40)
showing the value of \texttt{myObject.myProperty}.  When the user
clicks on the button, the \texttt{onClick} function (line 13) is
called. This function increases the value of
\texttt{myObject.myProperty} by one (line 15) and calls
\texttt{updateButton} function which updates the button's text to the new
value (line 22).  Once the page is loaded for the first time the
button shows \texttt{1} as the initial value of
\texttt{myObject.myProperty}.  In practice when the user clicks on the
button, \texttt{0} appears instead of \texttt{2}: there is a bug.

Two other functions are called in \texttt{onClick()}, \texttt{foo()}
and \texttt{bar()}. As developers we often encounter function calls
which seem peripheral to our current concern; they may have been added
by another developer, or we may have forgotten their exact properties
or those properties may have changed, and so on. The difference
between what we expect these functions to do, e.g. nothing
interesting, and what they do in practice may cause bugs.

@@

By browsing through the code or other means\cite{Barton},
 the developer determines that the value displayed on the 
button is set at line 22. Since the displayed value is incorrect we know 
the bug occurred before we hit this line.
To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, the execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is \texttt{0}. We expected this value to be \texttt{2}.

@@

As demonstrated in Figure~\ref{fig:example-points}, the developer has
examined three points of execution. The first point  was the breakpoint set by the developer. 
We call this special breakpoint the \textit{reproduction point}.
The second and third points preceded the reproduction point in execution sequence.
All three points-the history
of the search for the defect-are available through the debugger's
interface. On the top of the left panel in Figure~\ref{fig:example4}
there is an opened list which shows all three examined points. The
first one is the breakpoint on line 22, the second one is the point
which is when \texttt{myObject.myProperty} changed before
reaching the breakpoint and finally the last one is the point of
execution in which \texttt{myCondition.value} gets the
\texttt{undefined} value. Moreover, the source lines related to these
points are marked with red \textbf{Q} icons.

Notice that in our example, \textit{lastChange} combines some aspects
of breakpoint and of log-based debugging. Like breakpoint debugging,
the developer re-executes a live runtime without changing the source
and without a special execution environment beyond the debugger. The
state of the program memory and the call stack are available at each
lastChange point. Like log-based debugging, the program state and the
call stack are recorded during program execution. We can't halt the
program at \textit{lastChange} because we don't know which point is the last
one until we return to the original breakpoint. In section 5 we
discuss cases where it is possible to pause at lines of \textit{lastChange}.

\subsection{Re-Execution, Reproduction Point and Data Collection}

\section{Limitations and Future Work}

\section{Reproducible Non-Deterministic Execution}

\subsection{Combination of {\large\bf\textit{lastChange}} and Breakpoint Debugging}

\section{Conclusion}

\section{Acknowledgments}

@@

\bibitem{Barton}
J.J. Barton, and J. Odvarko. \newblock Dynamic and graphical web page breakpoints.
\newblock In \emph{Conference on World Wide Web(WWW)},
April, 2010.

@@

\bibitem{Horwitz}%[Hofer(2006)]
S. Horwitz, B. Liblit, and M. Polishchuk. \newblock Better Debugging via Output Tracing and Callstack-Sensitive Slicing.
\emph{IEEE Transactions on
Software Engineering}, 36(1):7-19, 2010.

\bibitem{JPDA}%[JSD(2010)]
Java Platform Debugger Architecture. \newblock http://java.sun.com/javase/technologies/ core/toolsapis/jpda.

@@

\bibitem{Korel}%[Ko(2008)]
B. Korel and J. Laski. \newblock Dynamic Program Slicing.
\newblock \emph{Information Processing Letters},
 29(3):155-163, 1988.

@@

\bibitem{Lienhard}%[Lienhard(2008)]
A. Lienhard, T. G\^{\i}rba, and O. Nierstrasz. \newblock Practical Object-Oriented Back-in-Time Debugging.
\newblock In \emph{22nd European conference on Object-Oriented Programming(ECOOP)},
July, 2008.

@@

\bibitem{JSD}%[JSD(2010)]
Mozila JavaScript Debugging Interface. \newblock http://www.mozilla.org/js/jsd.

@@

\bibitem{Sridharan}%[Boothe(2000)]
M. Sridharan, S.J. Fink , and R. Bodik. \newblock Thin slicing.
\newblock In \emph{Conference on Programming Language Design and Implementation(PLDI)},
June, 2007.

\bibitem{Weiser}
M. Weiser. \newblock Program slicing.
\newblock \emph{IEEE Transactions on
Software Engineering}, vol. SE-10, no. 4, pp. 352-357,
July 1984.

\bibitem{Zeller}
A. Zeller. \newblock Why programs fail: A guide to systematic debugging.
Morgan Kaufmann (2005)

@@

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
\end{document}

>>>>

\textit{Querypoint}, is a Firefox plugin which enhances the popular 
Firebug JavaScript debugger with a new, practical feature called 
\textit{lastChange}. \textit{lastChange} automatically locates the last point that 
a variable or an object property has been changed. Starting from a 
program halted on a breakpoint, the \textit{lastChange} algorithm 
applies queries to the live program during re-execution, recording 
the call stack and limited program state each time 
the property value changes. When the program halts again on the breakpoint, 
it shows the program state at the last change point. To evaluate the usability and effectivness 
of \textit{Querypoint} we studied four experienced JavaScript developers applying the tool to two test cases.

@@

%In breakpoint debugging, developers select data to be collected by
%searching through source files and setting breakpoints. To determine
%where a value was set incorrectly, a developer must set
%breakpoints at all possible points where the value changes. At
%every breakpoint, the developer must determine if the location is in
%fact related to the questionable value change then study the complex
%debugger user interface and memorize values or manually collect
%data. As the number of breakpoint hits increases, the process of
%checking the program state, collecting data and resuming the execution
%becomes cumbersome.

%In log-based debugging, developers select data to be collected by
%inserting statements for all points of possible change.  While in
%breakpoint-based debugging, the whole program state is available to
%developer, in log-based debugging, developer has to decide what data
%should be collected when inserts the log statement. It is very common
%that the developer has to repeat this step several times due to
%insufficient collected data, or to wait a long time because too much
%data is recorded. Once adequate data is collected, it still
%requires analyzing and understanding. Developers usually end up in
%dealing with long log files and analyzing huge amounts of collected data. 
%Neither approach effectively assists the developer in finding origins to a wrong
%value.

Locating the origin of wrong values becomes even harder when developers 
deal with weakly-typed dynamic languages such as JavaScript. First, due to weakly-typed
nature of these languages the search space that the developer should consider is considerably 
larger than traditional languages such as Java and C{\small\#}. Second, dynamic features make 
program full static analysis almost impossible.

\textit{Querypoint}, is a Firefox plugin which enhances the popular 
Firebug JavaScript debugger with a new, practical feature called 
\textit{lastChange}\cite{Mirghasemi2011}. \textit{lastChange}, locates the origin of a wrong
value by queries on the running program. This feature enables interactive 
dynamic data dependency navigation without requiring a full trace, unlike 
other dynamic slice navigation interfaces. The technique builds on existing 
breakpoint debugger technology  and it does not require a special environment to
create identical, instruction by instruction, re-executions. \textit{Querypoint} 
also provides mechanism for automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.

%Morover, it implements interactive 
% dynamic data dependency navigation for JavaScript, which poses a unique 
% challenge with respect to its dynamic nature.

%Imagine that a program execution 
%is paused on a breakpoint and the developer is
%suspicious about the value of a variable or an object property. The
%developer selects \textit{lastChange} on the value. The debugger
%replays the buggy execution and collects data when the data field changes. 
%Once the execution reaches the same place (i.e., the same
%breakpoint hit), it pauses the execution, analyzes the collected data
%and shows the location of the last change to the developer. The
%developer can also examine the program state at the located point of
%execution, and continue debugging by more \textit{lastChange} queries
%from that point.

\section{Related Work}
\label{sec:relatedWork}
%whyline, firecrystal
Most tools developed to enhance developers' navigation on the buggy execution can 
be classifed in two main groups: replay-based and logging-based. Replay-based 
approaches capture limited data during execution and replay the buggy execution 
to reach past points. In contrast, logging-based approaches collect enough data during
execution to relieve developer from re-execution, then query the data to 
inform the developer.

Among replay-based debuggers we can name bdb \cite{Boothe} and
reverse watchpoint \cite{Maruyama}. Both tools rely on deterministic
executions and employ a step counter to locate the requested point from the
beginning of execution. These tools incur two to four times runtime overhead.

Among logging-based debuggers are \textit{omniscient} debuggers (e.g.,
ODB\cite{Lewis}, TOD \cite{Pothier}, Unstuck\cite{Hofer} and WhyLine\cite{Ko}) and
time-travelling debuggers (e.g., Nirvana \cite{Bhansali}). Logging-based debuggers 
suffer from different issues. First, the recording
step is time expensive (20-120 times) and it should be repeated in case of changes in
program. Second, the execution log cannot fully replace the live
execution. There are other aspects of execution (e.g., program user
interface, file system, database tables, etc.) which are also
important in debugging and are not available to the developer in 
logging-based debuggers. Third, querying collected data (e.g., to restore
the program state at a certain point) may not be efficient enough for
debugging realistic programs.

\textit{Querypoint} resembles the operational model of replay-based debuggers
and the query approach of logging-based debuggers. Contrary to
other replay-based debuggers, which require exactly the same
re-executions (deterministic executions), \textit{Querypoint} only requires \textit{bug 
reproducibility}, meaning a test case is available which reproduces the bug and a 
way to halt execution reliably after the reproduction. Contrary to logging-based
debuggers \textit{Querypoint} selectively collects data which significantly reduces
runtime overhead incured by logging.

%A recent work by Lienhard et al.\cite{Lienhard} suggests virtual
%machine level support for keeping the history of events. It replaces every
%object reference with an alias object which keeps the history of
%changes to the object reference. Although this approach incurs less runtime overhead (7 times) in
%comparison to omniscient debuggers, it adds memory
%overhead.

% Origin tracking of \texttt{undefined} and \texttt{null} values employing \textit{value piggybacking} technique proposed by
%Bond et al. \cite{Bond}. This approach has two main limitations comparing to \textit{lastChange}.
%First, it is limited to \texttt{undefined} and \texttt{null} values. Second, it does not return the last change
%of a \texttt{null} variable but the first place that the \texttt{null} value is originated.

\section{Querypoint}
\label{sec:introExample}

We illustrate the \textit{Querypoint} functionality by a simple
example. The example demonstrates a buggy JavaScript code in a HTML
page (Figure~\ref{fig:js-code}). The page contains a button (line 40)
showing the value of \texttt{myObject.myProperty}.  When the user
clicks on the button, the \texttt{onClick} function (line 13) is
called. This function increases the value of
\texttt{myObject.myProperty} by one (line 15) and calls
\texttt{updateButton} function which updates the button's text to the new
value (line 22).  Once the page is loaded for the first time, the
button shows \texttt{1} as the initial value of
\texttt{myObject.myProperty}.  In practice when the user clicks on the
button, \texttt{0} appears instead of \texttt{2}: there is a bug.

%Two other functions are called in \texttt{onClick()}, \texttt{foo()}
%and \texttt{bar()}. As developers we often encounter function calls
%which seem peripheral to our current concern; they may have been added
%by another developer, or we may have forgotten their exact properties
%or those properties may have changed, and so on. The difference
%between what we expect these functions to do, e.g. nothing
%interesting, and what they do in practice may cause bugs.

@@

By browsing through the code,% or other means\cite{Barton},
 the developer determines that the value displayed on the 
button is set at line 22. Since the displayed value is incorrect we know 
the bug occurred before we hit this line.
To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, the execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is \texttt{0}. We expected this value to be \texttt{2}.

@@

The developer has examined three points of execution. The first point  was the breakpoint set by the developer. 
We call this special breakpoint the \textit{reproduction point}.
The second and third points preceded the reproduction point in execution sequence.
All three points-the history
of the search for the defect-are available through the debugger's
interface. On the top of the left panel in Figure~\ref{fig:example4}
there is an opened list which shows all three examined points. Moreover, the source lines related to these
points are marked with red \textbf{Q} icons.
%The
%first one is the breakpoint on line 22, the second one is the point
%which is when \texttt{myObject.myProperty} changed before
%reaching the breakpoint and finally the last one is the point of
%execution in which \texttt{myCondition.value} gets the
%\texttt{undefined} value.

%Notice that in our example, \textit{lastChange} combines some aspects
%of breakpoint and of log-based debugging. Like breakpoint debugging,
%the developer re-executes a live runtime without changing the source
%and without a special execution environment beyond the debugger. The
%state of the program memory and the call stack are available at each
%lastChange point. Like log-based debugging, the program state and the
%call stack are recorded during program execution. We can't halt the
%program at \textit{lastChange} because we don't know which point is the last
%one until we return to the original breakpoint. 
%In section 5 we discuss cases where it is possible to pause at lines of \textit{lastChange}.

\textit{Querypoint} needs a test case to reproduce the
execution and conditions to correctly recognize the reproduction point. 
Although both elements can be directly provided by developer, \textit{Querypoint}
is also able to automatically create them from the first execution.

To replay execution, \textit{Querypoint} keeps track of breakpoint hits and single steps. For example, 
if the developer queries \textit{lastChange} at the third hit of breakpoint \textit{b}, in
re-execution, the third hit is recognized as the reproduction point. \textit{Querypoint} supports two mechanism
for automatic re-execution: callstack-reproduction and record-replay. In callstack reproduction the function from
the earliest frame of the call stack is called with the same parameters. The record-replay execution uses two phases. In the record phase, it 
stores the initial page url and the events and parameters corresponding to user actions. In the replay phase, it opens
the same url and simulates events as if they were user actions.

In addition to the data collected at every change event for identifying the \textit{lastChange}
result, \textit{Querypoint} partially stores values in program state. There is a trade-off between the amount of data collected at every change event and the number of re-executions. If the developer asks for some values which have not been stored, \textit{Querypoint} re-executes and collects the requested data.

\section{User Study}
We supplied four experienced Javascript developers with \textit{Querypoint} in an 
extended Firebug debugger\footnote[1]{http://ltiwww.epfl.ch/\texttildelow mirghase/lastchange-userstudy}. Following a tutorial and a practice case, we observed as they 
applied both conventional breakpoint and \textit{lastChange} on two small programs we provided. The first program, Shapes, calculates the area and perimeter values for a list of shapes. The bug happens when one of the calculated numbers is zero. The second program, Moving Circle, randomly scales and moves a circle in the page. The bug happens once the circle becomes invisible after an exception occurs. This case represent a reproducible non-deterministic execution. The developers were asked to locate the defects that caused these bugs. All four developers successfully applied \textit{lastChange} to the test programs and understood how it could help debugging. To find the defect location with breakpoints, all four users took more steps\footnote[2]{A step is a button push, either single stepping the debugger or running a \textit{lastChange query}} and more time (Figure~\ref{fig:userstudy}).

%another searched the text, a third set a lot of breakpoints to understand the control flow. 
%Based on our own experience we expect these strategies represent the kinds of approaches 
%developers have available. These operations are time consuming and tedious. In contrast, all four users 
%found the defect location with just two \textit{lastChange} operations.

% Based on our measurements,
%this makes \textit{lastChange} about three times faster than breakpoints for the navigation from 
%the reproduction point to the defect
%We recognize that these programs were designed to highlight lastChange and many kinds of debugging 
%issues have been hidden by the design of our tests. Nevertheless our results show that, when a defect 
%relates to incorrect values and a developer recognizes this, then the operational 
%mechanics of lastChange lead to the defect much more quickly than breakpoints.

\begin{figure}[htp]
\centering 
\includegraphics[width=.48\textwidth]{10-userstudy.pdf}
\caption{The number of steps (and time in seconds) required before locating a defect, for each test subject and test program. 
Cells with a white background report values with conventional debugging; Cells with a colored background use \textit{lastChange}.}

\label{fig:userstudy}
\end{figure}

\section{Conclusion and Future Work}
\textit{Querypoint} provides critical information for debugging Jav- aScript programs: 
the location and state at the point where a questionable value was assigned. It only requires 
bug reproducibility and is built on the existing breakpoint technology.

In our next iteration we plan to merge the query and breakpoint results. \textit{Querypoint} shows the 
results of \textit{lastChange} in a similar but different view from breakpoint debugging. This focuses attention 
on value changes, but it makes studying control flow more difficult. %: we don't support single stepping from a \textit{lastChange} result in our user interface.

% In our next iteration we plan to merge the query and breakpoint results and support pausing as described in Sec.~\ref{sec:pausing}.

@@

%\bibitem{Barton}
%J.J. Barton, and J. Odvarko. \newblock Dynamic and graphical web page breakpoints.
%\newblock In \emph{Conference on World Wide Web(WWW)},
%April, 2010.

@@

%\bibitem{Horwitz}%[Hofer(2006)]
%S. Horwitz, B. Liblit, and M. Polishchuk. \newblock Better Debugging via Output Tracing and Callstack-Sensitive Slicing.
%\emph{IEEE Transactions on
%Software Engineering}, 36(1):7-19, 2010.

%\bibitem{JPDA}%[JSD(2010)]
%Java Platform Debugger Architecture. \newblock http://java.sun.com/javase/technologies/ core/toolsapis/jpda.

@@

%\bibitem{Korel}%[Ko(2008)]
%B. Korel and J. Laski. \newblock Dynamic Program Slicing.
%\newblock \emph{Information Processing Letters},
% 29(3):155-163, 1988.

@@

%\bibitem{Lienhard}%[Lienhard(2008)]
%A. Lienhard, T. G\^{\i}rba, and O. Nierstrasz. \newblock Practical Object-Oriented Back-in-Time Debugging.
%\newblock In \emph{22nd European conference on Object-Oriented Programming(ECOOP)},
%July, 2008.

@@

\bibitem{Mirghasemi2011}
S. Mirghasemi, J.J. Barton, and C. Petitpierre. \newblock Debugging by lastChange.
\newblock Technical Report. EPFL-REPORT-164250, 2011. 
%\bibitem{JSD}%[JSD(2010)]
%Mozila JavaScript Debugging Interface. \newblock http://www.mozilla.org/js/jsd.

@@

%\bibitem{Sridharan}%[Boothe(2000)]
%M. Sridharan, S.J. Fink , and R. Bodik. \newblock Thin slicing.
%\newblock In \emph{Conference on Programming Language Design and Implementation(PLDI)},
%June, 2007.

%\bibitem{Weiser}
%M. Weiser. \newblock Program slicing.
%\newblock \emph{IEEE Transactions on
%Software Engineering}, vol. SE-10, no. 4, pp. 352-357,
%July 1984.

%\bibitem{Zeller}
%A. Zeller. \newblock Why programs fail: A guide to systematic debugging.
%Morgan Kaufmann (2005)

@@

% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%

\appendix
\section{Presentation}
The presentation will include the following parts:
\begin{enumerate}
	\item A short demo, debugging a simple buggy web page by the tool.
	\item The brief explanation of \textit{lastChange} algorithm.
	\item The brief explanation of \textit{Querypoint} debugging as a general debugging approach.
	\item The Discussion about non-deterministic reproducible executions.
	\item A Comprehensive demo on the Tool features.
\end{enumerate}

\section{Tool Maturity}
The tool is functional, however this is the first version of \textit{Querypoint} and therefore it is 
not mature enough to be used by developers. Moreover, \textit{Querypoint} is built on top of the 
Firefox 3.6 JavaScript engine. Due to significant changes in Firefox platform, \textit{Querypoint} is 
not compatible with the newer versions of Firefox.

We requested the mozilla team to support basic operations needed by \textit{lastChange} inside the 
Firefox JavaScript Engine. Our proposal was accepted and they planned for the implementation of 
these operations (Feature requests 641234 and 641236 in Mozilla Bugzilla) in the next major Firefox release.

\section{Resources}
The tool and its source code are publicly available and can be downloaded 
from \textit{Querypoint Debugging} web site at this address:

\mbox{\href{http://code.google.com/p/querypoint-debugging}{http://code.google.com/p/querypoint-debugging}}

The details of our user study, including the instructions and test cases source code, are available at this address:

\mbox{\href{http://ltiwww.epfl.ch/~mirghase/lastchange-userstudy}{http://ltiwww.epfl.ch/\texttildelow mirghase/lastchange-userstudy}}

\end{document}
